Emerging role of oncogenic ß-catenin in exosome biogenesis as a driver of immune escape in hepatocellular carcinoma

Version 1

Abstract

Abstract

Immune checkpoint inhibitors have produced encouraging results in cancer patients. However, the majority of ß-catenin mutated tumors have been described as lacking immune infiltrates and resistant to immunotherapy. The mechanisms by which the oncogenic ß-catenin affects immune surveillance remain unclear. Herein, we highlighted the involvement of ß-catenin in the regulation of the exosomal pathway and in the immune/cancer cell communication in hepatocellular carcinoma (HCC). We showed that mutated ß-catenin represses SDC4 and RAB27A expression, two main actors of exosome biogenesis, in both liver cancer cell lines and HCC patient samples. Using nanoparticle tracking analysis and live-cell imaging, we further demonstrated that activated ß-catenin represses exosome release. Then, we demonstrated in 3D spheroid models that activation of β-catenin promotes a decrease of immune cell infiltration through a default in exosome secretion. Taken together, our results provide the first evidence that oncogenic ß-catenin plays a key role in exosome biogenesis. Our study gives new insight into the impact of ß-catenin mutations on tumor microenvironment remodeling, which could lead to the development of new strategies to enhance immunotherapy response.

Introduction

ß-catenin, encoded by the CTNNB1 gene, is a main oncogene, mutated in various cancers, such as liver, endometrial, colorectal cancers or melanoma (1). These ß-catenin-mutated tumors share specific features such as a microenvironment devoid of immune infiltrates. Thus, in front of the therapeutic revolution for cancers, with the emergence of immunotherapy and more particularly immune checkpoint inhibitors (anti-PD1, anti-PD-L1), most of the ß-catenin mutated tumors remain resistant to immunotherapy (2–9). In these tumors, the oncogenic β-catenin is able to set up a microenvironment that favors tumor progression notably by promoting immune escape. Few studies have reported that the defect of intercellular communication between cancer cells and immune cells is partly mediated by a decrease of chemokines production leading to a reduction of immune infiltrates. In melanoma and hepatocellular carcinoma (HCC), a decrease of, respectively CCL4 and CCL5, leads to a defective recruitment of dendritic cells and consequently impaired T-cell activity and immune escape (2,5). As immunotherapies are now part of the therapeutic arsenal for cancers, it is important to better understand the mechanisms underlying this immune escape phenotype. In this study we focused on ß-catenin-mutated HCC and intercellular communication through extracellular vesicles.

HCC is the most frequent form of primary liver adult cancer (80% of cases). HCC represents the 6th most common diagnosed cancer worldwide and the 3rd leading cause of cancer-related death (10). This pathology has a poor prognosis and the diagnosis is often made at an advanced stage for which treatment options are limited, especially if surgery is no longer possible. In 2020, the combination of Atezolizumab (anti-PD-L1) and Bevacizumab (anti-VEGF) became the first-line FDA-approved therapy for advanced HCC (11). However, as described previously, β-catenin-mutated HCC are characterized by immune escape and are resistant to immunotherapy (2,3,12,13). CTNNB1 gene mutations are found in 30-40% of HCC and trigger uncontrolled transcriptional activity (14,15). These mutations prevent β-catenin degradation, fostering its role as a transcriptional co-factor and thus the expression of genes involved in cell proliferation and invasion. The mechanistic bases underlying immune evasion in β-catenin mutated HCC are not yet fully understood.

Tumor-derived extracellular vesicles (EVs) are increasingly described as important actors of the tumor microenvironment communication (16); but no studies explored their role in the immune escape and immunotherapy resistance of ß-catenin mutated tumors. EVs are nanometric phospholipid bilayer vesicles, which contain various cargoes such as RNAs or proteins and that impact recipient cells properties and functions. EVs can be classified into two main categories, namely, microvesicles and exosomes. Microvesicles (50 to 1000 nm) correspond to EVs directly secreted from the plasma membrane (PM). Whereas exosomes (50 to 150 nm) are released from an endosomal compartment, the multivesicular body (MVB), by its fusion with the cell surface (16). Among the actors implicated in exosome biogenesis and secretion, syndecan-4, involved in the endosomal membrane budding (17) and Rab27a which promotes the docking of MVB to the PM play key roles in exosome release (18). As important mediators of cell-to-cell communication, EVs can regulate tumor growth, angiogenesis, invasion and infiltration of immune cells into the tumor (19). In the context of HCC, it was reported that EVs derived from tumor cells can modulate epithelial-to-mesenchymal transition, intrahepatic metastasis and tumor immunity (20).

Based on these considerations, the current study investigated the role of exosomes in mutated ß-catenin mediated immune evasion in HCC. We demonstrated that mutated ß-catenin represses syndecan-4 and Rab27a expression, thus altering exosomal secretion from HCC cells, leading to a defective recruitment of immune cells in tumors.

Results

Silencing of mutated ß-catenin increases exosome biogenesis-associated gene expression in HepG2 cells

CTNNB1 mutations in HCC are often monoallelic, leaving a wild-type allele in tumor cells. To specifically study the oncogenic ß-catenin, we made use of the dual ß-catenin knockdown (KD) HepG2 model we published recently (21). The transcriptional analysis of HepG2 KD cells revealed that the expression of 1973 genes (log2(Fold-change) >0.5 and <-0.5) is modulated upon mutated ß-catenin silencing (21) (Figure 1a). Gene ontology analysis (FunRich) performed on upregulated genes showed that mutated ß-catenin silencing enhanced significantly the expression of genes linked to specific cellular components including exosomes (Figure 1b). Moreover, silencing of the mutated ß-catenin led to the overexpression of 17 genes associated to exosome biogenesis such as RAB27A and SDC4 (Figure 1c). Similarly, proteomic analysis of HepG2 KD cells revealed exosomes as the main cellular component displaying up-regulated proteins upon mutated ß-catenin silencing (Figures 1d-1e). Among these 9 up-regulated proteins we found Rab27a and Syndecan-4, consistent with the transcriptomic analysis results (Figure 1f). These results suggest the involvement of mutated ß-catenin in the regulation of the exosomal pathway.

Mutated ß-catenin regulates exosome biogenesis gene expression in HepG2 cells.

(a-f) HepG2 cells were transfected with siRNAs targeting mutated ß-catenin or control siRNAs. (a) Volcano plot of deregulated genes identified by microarray based-transcriptomic analysis. Red and blue dots indicated respectively significantly up- and down-regulated genes. (b) Upregulation of cellular component genes using FunRich software. (c) Upregulated genes associated with exosome biogenesis pathway. The graph indicates the fold change (FC) when comparing the mutated ß-catenin silencing condition with the control condition. (d) Volcano plot of deregulated proteins identified by mass spectrometry. Red and blue dots indicated respectively significantly up- and down-regulated proteins. (e) Upregulation of cellular component proteins identified using FunRich software. (f) Upregulated proteins associated with exosome biogenesis pathway. The graph indicates the fold change when comparing the mutated ß-catenin silencing condition with the control condition. Results are expressed as Mean ± SEM, two-tailed Student’s t-test analysis. *P<0.05; **P<0.01; ***P<0.001; ****P<0.0001.

Stable depletion of mutated ß-catenin increases exosome secretion in the HepG2 cell model

In the model presented above, siRNAs were used to transiently silence mutated ß-catenin expression in HepG2 cells (21). To improve this method, the same sequences were used to develop an inducible shRNA strategy to stably inhibit the expression of the mutated ß-catenin. The use of doxycycline inducible promoters allowed regulating the expression of shRNAs over time for more in-depth and specific investigations of ß-catenin function in various cellular processes. In HepG2 cells, this strategy reduced protein expression by 62% for mutated ß-catenin and by 42% for CyclinD1, a ß-catenin positive transcriptional target (Figure 2a). The model was also validated regarding mRNA expression of several ß-catenin transcriptional targets with a decrease of CCND1 and AXIN2 expression (positive targets) and an increase in ARG1 expression (negative target) (Figure sup 1a). Also, as we previously described with the siRNA strategy (21), after mutated ß-catenin depletion, we observed an increase in the number and in the size of bile canaliculi (BC), feature of more differentiated cells, in HepG2 cells (Figures sup 1b-1c). Nanoparticle-tracking analysis (NTA) revealed that the depletion of the mutated ß-catenin by both siRNA and shRNA strategies increased the number of secreted particles (Figure 2b, Figure sup 1d). Using NTA, the particle mean size, which was not altered between both conditions, was around 100-150 nm, compatible with the size of small EVs such as exosomes (Figure 2b, Figure sup 1d). To analyze the exosomal release in living HepG2 cells, we used a pH-sensitive reporter (CD63-pHluorin) allowing the visualization of MVB–PM fusion by total internal reflection fluorescence (TIRF) microscopy (22). The tetraspanin CD63 is a well-known exosome surface marker (23,24). In cells expressing CD63-pHluorin, we could detect multiple increases in the fluorescence signal at the PM over the time suggesting fusion of CD63-pHluorin– positive acidic vesicles with the PM (Figure sup 1e). We found that mutated ß-catenin depletion in HepG2 cells significantly increased the number of fusion events, corresponding to secreted exosomes (Figure 2c). To go further, EVs released by HepG2 cells were isolated by differential ultracentrifugation and characterized (Figure 2d). First, western-blot analysis of isolated EVs confirmed the expression of the CD63 exosomal marker. It is noteworthy that we observed an increase of CD63 expression after mutated ß-catenin depletion suggesting an increase of the exosomal fraction (Figure 2e). This increased expression of CD63 is also confirmed with the proteomic analysis of HepG2 KD cells (Figure 1f). Moreover, the WT and the mutated forms of the ß-catenin protein were also found in HepG2-derived EVs and the expression of mutated ß-catenin is decreased according to the cells of origin (Figure 2e). Then by transmission electron microscopy (TEM), we demonstrated that HepG2-derived EVs are similar to exosomes regarding their morphology (flat cup-shaped) and their size (50-150nm) (25) (Figures 2f and sup 1f). To better understand the regulation of exosome secretion by mutated ß-catenin, we performed TEM experiments in order to visualize MVBs, the origin of exosomes. TEM analysis revealed a higher number of MVBs per cell and an increase of their diameter upon mutated ß-catenin removal (Figure 2g). Taken together these results highlight that mutated ß-catenin represses exosome secretion.

Mutated ß-catenin controls exosome secretion in HepG2 cells.

(a-g) HepG2-shßCat MUT and HepG2-shCtrl cells were treated with doxycycline (DOX) to silence or not mutated ß-catenin. (a) Analysis of ß-catenin and CyclinD1 expression by western-blot. Stain free was used as loading control. Graphs show the quantification of seven independent experiments. (b) Nanoparticle tracking analysis of supernatant. Graphs show the quantification of seven independent experiments. (c) Quantification of CD63-pHluorin MVB–PM fusion events visualized by live TIRF microscopy. Depicted data are representative of three independent experiments, each dot represents one cell. Images represent the cell mask (white) and red dots corresponding to fusion events. Scale bar: 10µm. (d) Extracellular vesicles (EVs) isolation protocol. Created with BioRender.com. (e) Analysis of ß-catenin and CD63 expression in HepG2-derived EVs. Stain free was used as loading control. Graphs show the quantification of four independent experiments. (f) Transmission electron microscopy images of HepG2-derived EVs by close-up. Scale bar: 100nm. The graph shows the diameter quantification of EVs (n=93). (g) Electron microscopy images of HepG2 shCtrl and shßcat MUT cells showing multivesicular bodies (MVBs) (yellow arrowheads). Scale bar: 2 µm (zoom: 500 nm). The graphs show the quantification of the number of MVBs per cell and the MVB diameter. (a-f) Results are expressed as Mean ± SEM, two-tailed Student’s t-test analysis. *P<0.05; **P<0.01; ***P<0.001; ns, non-significant.

Mutated ß-catenin decreases Syndecan-4 and Rab27a expression in liver cancer cell lines

To explore the molecular aspect of the alteration of exosome secretion, we first attempt to confirm the data obtained in our transcriptomic and proteomic approaches. In HepG2 cells, we confirmed that transient (Figures sup 2a-b) and stable (Figures 3a-3b) mutated ß-catenin depletion increased both SDC4 and RAB27A gene and protein expression. Due to the lack of efficient antibodies against SDC4, we failed in analyzing SDC4 protein expression by Western-Blot. We confirmed the increase of expression of Syndecan-4 and Rab27a by immunofluorescence (Figures 3c-3d). We then aimed to extend these results to other liver cancer cell lines. We analyzed SDC4, RAB27A, ARG1 and AXIN2 basal mRNA expression in five different liver cancer cell lines, that carry a different CTNNB1 mutational status, and identified positive correlations between SDC4 and RAB27A with ARG1 (ß-catenin negative target) and negative correlations between SDC4 and RAB27A with AXIN2 (ß-catenin positive target) (Figure sup 2c-d). Then, we analyzed Rab27a basal protein expression in these five cell lines, and identified a negative correlation between ß-catenin and Rab27a expression (Figures sup 2f-2g). In the two liver cancer cell lines bearing a ß-catenin point mutation (Huh6 and SNU398 cells), inducible shRNA strategy with doxycycline respectively reduced by 63% and 49% the ß-catenin protein expression (Figures sup 2h-2i). We validated both Huh6 and SNU398 models using the shRNA ß-catenin strategy as we observed a decrease of CCND1 and AXIN2 mRNA expression and an increase of ARG1 mRNA expression (Figures sup 2j-2k). In agreement with the results obtained in HepG2 cells, ß-catenin depletion also increased SDC4 and RAB27A gene expression and Rab27a protein expression in Huh6 and SNU398 cells treated with doxycycline (Figures 3e-h). In order to mimic the ß-catenin activation in a non-mutated HCC cell line, we treated Huh7 cells with a GSK3 inhibitor (CHIR99021) limiting ß-catenin phosphorylation and degradation. We demonstrated that CHIR99021 treatment increased ß-catenin protein expression (2.37-fold) (Figure 3i) and induced its translocation into the nucleus in Huh7 cells (Figure 3j). The presence of ß-catenin in the nucleus was associated with the deregulation of ß-catenin transcriptional targets (Figure sup 2l). Consistent with our data, CHIR99021 treatment decreased SDC4 and RAB27A gene expression, and Rab27a protein level in Huh7 cells (Figures 3k-l). Then, we analyzed the exosomal release by TIRF microscopy in living Huh7 cells expressing CD63-pHluorin. We found that CHIR99021 treatment significantly decreased the number of fusion events in Huh7 cells (Figure 3m). All together, these results revealed that ß-catenin mutation/activation represses the expression of syndecan-4 and Rab27A, two proteins involved in exosome biogenesis, associated with a repression of exosomal secretion.

Activated ß-catenin represses SDC4 and RAB27A expression in liver cancer cells.

(a-d) HepG2 cells were treated with doxycycline (DOX) to express either a control shRNA (shCtrl) or a shRNA targeting mutated ß-catenin (shßcat MUT). (a) Analysis of SDC4 and RAB27A mRNA expression by qRT-PCR. Graphs show the quantification. (b) Analysis of Rab27a protein expression by western-blot. Stain free was used as loading control. The graph shows the quantification. (c-d) Epifluorescence images of HepG2 shCtrl and shßcat MUT cells stained with Syndecan-4 and Rab27a antibodies (red), Phalloidin (green), Hoechst (blue). Scale bar: 20 µm. Graphs show the quantification of the fluorescence intensity per image and divided per nuclei. Depicted data are representative of three independent experiments, each dot represents one image. (e) Analysis of SDC4 and RAB27A mRNA expression in Huh6 cells expressing either a control shRNA (shCtrl) or a shRNA targeting ß-catenin (shßcat) treated with DOX. The graph shows the quantification. (f) Analysis of SDC4 and RAB27A mRNA expression in SNU398 cells expressing either a control shRNA (shCtrl) or a shRNA targeting ß-catenin (shßcat) treated with DOX. The graph shows the quantification. (g-h) Analysis of Rab27a protein expression in Huh6 (g) or SNU398 (h) cells expressing either a control shRNA (shCtrl) or a shRNA targeting ß-catenin (shßcat). Stain free was used as loading control. The graphs show the quantification. (i-m) Huh7 cells treated with DMSO or CHIR99021 (3µM) for 48 hours. (i) Analysis of ß-catenin expression by western-blot. The graph shows the quantification. (j) Epifluorescence images of cells stained with ß-catenin antibody (red), Phalloidin (green), Hoechst (blue). Scale bar: 20 µm. (k) Analysis of SDC4 and RAB27A mRNA expression by qRT-PCR. The graphs show the quantification. (l) Analysis of Rab27a protein expression by western-blot. The graph shows the quantification. (m) Quantification of CD63-pHluorin MVB–PM fusion events visualized by live TIRF microscopy. Depicted data are representative of three independent experiments; each dot represents one cell. Images represent the cell mask (white) and red dots corresponding to fusion events. (a-m) All graphs show the quantification of at least three independent experiments. Results are expressed as Mean ± SEM, two-tailed Student’s t-test analysis. *P<0.05; **P<0.01; ***P<0.001.

Mutated ß-catenin promotes immune evasion in HCC cells through exosomes

We aim at deciphering the involvement of mutated ß-catenin in HCC immune escape through the regulation of exosomes. We used 3D spheroid models from two liver cancer cell lines (HepG2 and Huh7) and after 24h co-culture with Peripheral Blood Mononuclear Cells (PBMC), immune cell infiltration was evaluated by flow cytometry (Figures 4a-4b). First of all, we assessed the capability of PBMC to invade spheroids formed from Huh7 cells treated with CHIR9901. We observed a decrease of PBMC infiltration and an increase of tumor cell survival when Huh7 cells were treated with CHIR9901 compare to control condition (Figure 4c). Of note, we confirmed that CHIR99021 increased ß-catenin and decreased Rab27a protein expression in Huh7 spheroids with no impact on the cell viability (Figures sup 3a-3b). Next, we showed that the depletion of mutated ß-catenin protein enhanced PBMC infiltration in HepG2 cells associated with a decrease of tumor cell survival (Figure 4d). We validated that mutated ß-catenin protein is correctly depleted in HepG2 spheroids, with an increase of Rab27a protein expression and with no impact on the cell viability (Figures sup 3c-d). Altogether, these results confirm the involvement of mutated ß-catenin in immune evasion. We next sought to investigate the role of exosomes in this process. In order to blunt exosome secretion, we used siRNA directed against Rab27a in spheroids formed with HepG2 cells depleted for the mutated ß-catenin. Rab27a depletion decreases the number of secreted particles without any effect on cell viability in HepG2 shßcat MUT spheroids (Figures sup 3e-g). Furthermore, depletion of Rab27a decreased PBMC infiltration associated with an increase of tumor cell survival (Figure 4e), strongly suggesting the determinant role of exosomes in immune cell infiltration.

Activated ß-catenin represses immune infiltration in liver cancer spheroids through exosomes.

(a) Peripheral Blood Mononuclear Cells (PBMC) infiltration analysis protocol. Created with BioRender.com. (b) Images of HepG2 spheroids expressing a control shRNA incubated or not with PBMC for 24 hours. (c) Analysis of PBMC infiltration and tumor cell survival in Huh7 spheroids treated with DMSO or CHIR99021. (d) Analysis of PBMC infiltration and tumor cell survival in HepG2 spheroids expressing control shRNA (shCtrl) or shRNA targeting mutated ß-catenin (shßcat MUT). (e) Analysis of PBMC infiltration and tumor cell survival in HepG2 spheroids co-expressing shRNA targeting mutated ß-catenin (shßcat MUT) and control siRNA (siCtrl) or siRNA targeting Rab27A (siRab27A). (c-e) Graphs show the quantification of six independent experiments. Results are expressed as Mean ± SEM, one or two-tailed Student’s t-test analysis. *P<0.05; **P<0.01; ***P<0.001.

CTNNB1 mutations are associated to low expression of exosomal biogenesis-associated genes in HCC patient samples

To determine whether markers of exosomal biogenesis are deregulated in HCC human samples, we assessed the expression of SDC4 and RAB27A using public transcriptomic data from two different cohorts (Figure 5a: TGCA data sets from Cbioportal, Figure 5b: Boyault et al. article (26)). In both set of data, we showed that SDC4 and RAB27A gene expressions were significantly reduced in CTNNB1-mutated HCCs compare to non-mutated tumors (Figures 5a-5b). Furthermore, Pearson’s correlation analysis of these two cohorts revealed that RAB27A and SDC4 gene expression were positively correlated with the expression of ARG1 and PCK1 (ß-catenin negative targets) (27) and negatively correlated with the expression of GLUL, CCND1, AXIN2, LGR5, FAT1 and BMP4 (ß-catenin positive targets) (Figures sup 4a-b). We then performed immunohistochemistry analysis on 56 human HCC samples bearing or not CTNNB1 mutations (Figure sup 4c). We found that tumors strongly positive for glutamine synthetase (a well-known ß-catenin positive target in the liver (28)), were correlated to low levels of Rab27a protein expression (Figure 5c), supporting the link between ß-catenin activation and regulation of exosomal secretion through a decrease of Rab27a expression.

Downregulation of SDC4 and RAB27A in human HCC samples with CTNNB1 mutations.

(a) Analysis of SDC4 and RAB27A mRNA expression in ß-catenin mutated (red) and non-mutated (blue) HCC (Cbioportal cohort, n=366). (b) Analysis of SDC4 and RAB27A mRNA expression in ß-catenin mutated (red) and non-mutated (blue) HCC (Boyault et al. cohort, n=56). (c) Immunohistochemistry (IHC) analysis of glutamine synthetase (GS) and Rab27a in HCC samples presenting or not ß-catenin mutations. Scale bar: 1mm. (T: tumoral, NT: non tumoral). The upper graphs show the quantification of the percentage of cells positive for GS or Rab27A. The analysis was split into two cohorts (circle and triangle). The lower graph shows the Pearson correlation between Rab27a and GS protein expression (n=56). Results are expressed as Mean ± SEM, two-tailed Student’s t-test analysis or Pearson correlation test. *P<0.05; **P<0.01; ***P<0.001; ****P<0.0001.

Discussion

ß-catenin is a main oncogene in tumors (1) and several studies have reported a key role for mutated ß-catenin in immune escape in several types of cancer such as HCC, colorectal cancer, melanoma and glioblastoma (2–9) but its mechanisms of action in this process remain to be explored deeper. During the past decades, ß-catenin mutated HCC have been classified at the molecular, histo-pathological and clinical levels as liver cancers with specific features (29–31). More recently, these tumors have been described as cold with low immune infiltrates and resistant to immunotherapy (2,3,12,13).

In this study we deciphered the role of mutated ß-catenin in cell communication between cancer cells and immune cells. Based on omic results, we hypothesized a role of EV in the control of immune infiltrates in ß-catenin mutated HCC. Following ISEV guidelines (32), we applied various techniques to address the impact of ß-catenin activation on EV secretion. Using NTA, we demonstrated that HepG2 cells silenced for mutated ß-catenin secrete more particles than control cells. Our different analyses and results (TIRF with the CD63 exosomal marker, TEM for the EVs shape and size, regulation of Rab27a expression) led us to take a strong interest in small EVs, i.e. exosomes, a subpopulation of EVs. Using real-time visualization of the release of CD63-enriched EVs we demonstrated that mutated ß-catenin silencing in HepG2 cells favors small EV secretion. The link between activation of ß-catenin and secretion of small EVs was further confirmed in ß-catenin non-mutated cells, Huh7 cells, treated by GSK3 inhibitor, suggesting that activation of the Wnt/ß-catenin pathway may result in a repression of the EV machinery in tumor cells. This link between ß-catenin and EVs was poorly explored in the literature and essentially restricted to the identification of ß-catenin as a common EV cargo (33). For example, it was shown that tetraspanins expression reduces the cellular pool of β-catenin by enhancing the exosome-associated export of β-catenin from the cell and that this exosomal discharge of β-catenin downregulates the Wnt signaling pathway (34). More recently, using a genome-wide CRISPR/Cas9-mediated screening to identify genes involved in EV release, Lu et al. identified the Wnt signaling pathway in the top hits (35). Indeed, by this approach, a number of Wnt-signaling modulators appeared among the candidates as EV biogenesis regulators, which were further validated in chronic myelogenous leukemia K562 cells. They also demonstrated that Wnt-mediated GSK3 inactivation, in K562 cells, regulated EV release in a Rab27a dependent manner (35). Herein, we added new findings reporting that this regulation also occurs in solid tumors. Moreover, we demonstrated for the first time the specific involvement of the oncogenic ß-catenin in the regulation of exosome biogenesis.

At the molecular level, thanks to global approaches we found that oncogenic ß-catenin represses the expression of genes involved in the EV machinery. We indeed found that genes from the Rab (RAB8B, RAB11A, RAB11B, RAB17, RAB27A), Vamp (VAMP2, VAMP5, VAMP8) and Vps (VPS4A, VPS4B, VPS16, VPS28) families were upregulated upon removal of the oncogenic ß-catenin in HepG2 cells. Herein, we focused on RAB27A and SDC4 genes, which were the most dysregulated in our transcriptomic analysis. We showed that mutated ß-catenin represses their gene expression in liver cancer cell lines. In non-mutated ß-catenin HCC cells, activation of ß-catenin showed that both gene and protein expression of Rab27a were decreased.

We further found a correlation between the presence of CTNNB1 mutations and the low expression of RAB27A and SDC4 in human HCC patient samples. The way RAB27A and SDC4 gene expression is repressed by ß-catenin remains to be explored. We described positive correlation between expression of SDC4 and RAB27A and expression of negative transcriptional targets of ß-catenin, and negative correlation with expression of positive targets in tumors. Even if the mechanism by which ß-catenin and LEF/TCF transcription factors led to target gene expression activation is well known (21,36,37), it is not the case for the repression process. In melanoma the transcription repressor ATF3 induced by the Wnt/ß-catenin pathway has been associated with CCL4 repression (5). Analysis of SDC4 and RAB27A promoter sequences revealed the presence of consensus sequences of Lef1, Tcf4 and Atf3 binding (data not shown), suggesting that these newly identified target genes could be regulated at the transcriptional level by transcription factors known to associate with ß-catenin. In this line, SDC4 was found as a negative Tcf4 target gene in HCC mouse models (38).

Following the EV machinery alteration, we further addressed the impact of the ß-catenin-mediated repression of EV secretion on immune cell infiltration in tumors. Herein, the functional role of these tumor-derived exosomes in tumor immunity was studied using liver cancer 3D models and PBMC. We showed that mutated ß-catenin promotes a decrease of immune cell infiltration in micro-tumors through a decrease of exosome secretion by depletion of Rab27a using siRNA. The small GTPase Rab27a has been described in several cancer progression mechanisms such as metastasis, cell invasion or migration (39). The oncogenic function of Rab27a is mainly due to its function as regulator of exosome secretion, which modulates cancer cell function and tumor microenvironment. In mammary carcinoma cells, the decrease of exosome secretion by Rab27a inhibition resulted in a diminution of primary tumor growth and lung metastatic dissemination (40). However, shRNA knockdown of Rab27a in a prostate cancer model revealed a dominant immunosuppressive role for EVs during antigen cross presentation (41). These data are inconsistent with our own, meaning that Rab27a-induced exosomal secretion is a double-edged sword in tumor progression. This could be due to the content of EVs. In fact, EVs are known to contain heterogeneous cargos that can play a critical role in immunomodulation (42). For example, the main immunosuppressive cytokine TGF-ß expressed by tumoral-derived EVs has been identified in Natural Killer cell function suppression (43,44). EVs released by tumoral cells can also carry several miRNAs mainly known to be involved in immune suppression (42). MiR-103a and Let-7a in tumor EVs promote M2-like polarization in lung and melanoma cancers (45,46). Interestingly, immunomodulatory proteins such as CSF-1, CCL2, FTH, FTL, and TGF-β chemokines have been identified in tumor exosomes (46). The content of ß-catenin-regulated EVs remains to be explored to fully understand their function in the immunomodulation of the tumor microenvironment.

In conclusion, this study is the first description of the oncogenic ß-catenin as a key regulator of the exosome machinery. Moreover, it gave new insights regarding the regulation of tumor-derived exosome production by the ß-catenin pathway and their involvement in the immune escape observed in ß-catenin mutated HCC. Given that ß-catenin is also mutated in melanoma, endometrial and colorectal cancers, these new findings may have a broader significance. Our data indicate that exosomes could be a promising biomarker and that restoring the tumor exosomal secretion of ß-catenin-mutated tumors may increase the communication between tumor and immune cells, and turning hot, cold tumors.

Supporting information

eLife Assessment

This study reports compelling data supporting the role of beta-catenin on intercellular communication occurring via extracellular vesicles, with implications for immune evasion in hepatocellular carcinoma (HCC). This fundamental insight sheds light on the underlying biology of HCC at a time when an increasing number of treatment options, including targeted small molecular inhibitors and immuno-oncologic drugs are being used to treat patients in the absence of validated predictive biomarkers. This work will be of special interest to researchers investigating the basic biology of liver function and HCC, and also to readers investigating novel pathways for therapeutic targeting of HCC.

Reviewer #1 (Public Review):

Summary:

This finding shows a connection between cancer-associated beta-catenin mutations and extracellular vesicle secretion. A link between the beta-catenin mutation and expression of trafficking and exocytosis machinery. They used a multidisciplinary approach to explore expression levels of relevant proteins and single-particle imaging to directly explore the release of extracellular vesicles. These results suggest a role of extracellular vesicles in immune evasion in liver cancer with the role needing to be further explored in other forms of cancer. I find this work to be compelling and of strong significance.

Strengths:

This paper uses multidisciplinary methods to demonstrate the compelling role of beta-catenin mutations in suppressing EV secretion in tumors. The results and imaging are extremely convincing and compelling.

Weaknesses:

There is no major weakness in this work. There are only things that left me more intrigued about this work. While the role of Rab27 was strongly examined, the hits of the VAMP proteins were not explored in detail. I was wondering if the decrease in the presence of VAMPS directly suggests the final step of membrane fusion in the exocytosis of EVs is what is being impaired. Or if it is other trafficking steps along the EV secretion pathway.

Reviewer #2 (Public Review):

Summary:

Dantzer and colleagues are investigating the pivotal role of ß-catenin, a gene that undergoes mutation in various cancer cells, and its influence on promoting the evasion of immune cells. In their initial experiments, the authors developed a HepG2 mutated ß-catenin KD model, conducting transcriptional and proteomic analyses. The results revealed that the silencing of mutated ß-catenin in HepG2 cells led to an up-regulation in the expression of exosome biogenesis genes.

Furthermore, the researchers verified that these KD cells exhibited increased production of exosomes, with the mutant form of ß-catenin concurrently decreasing the expression of SDC4 and Rab27a. Intriguingly, applying a GSK inhibitor to the cells resulted in reduced expression of SDC4 and Rab27a. Subsequent findings indicated that mutated ß-catenin actively facilitates immune escape through exosomes, and silencing exosome biogenesis correlates with a decrease in immune cell infiltration.

In a crucial clinical correlation, the study demonstrated that patients with ß-catenin mutations exhibited low levels of exosome biogenesis.

Strengths:

Overall, the data robustly supports the outlined conclusions, and the study is commendably designed and executed. However, there are a few suggestions for manuscript improvement.

Weaknesses:

No weaknesses were identified by this reviewer.

Reviewer #3 (Public Review):

Summary:

In this very important study by Dantzer et al., 'Emerging role of oncogenic b-catenin in exosome biogenesis as a driver of immune escape in hepatocellular carcinoma' the authors define a role for oncogenic b-catenin on exosome biology and explore the link between reduce exosome secretion and tumor immune cell evasion. Using transcriptional and proteomic analysis of hepatocellular carcinoma cells with either oncogenic or wildtype b-catenin the authors find that oncogenic b-catenin negatively regulates exosome biogenesis.

The authors can provide compelling evidence that oncogenic b-catenin in different hepatocellular carcinoma cells negatively regulates exosome biogenesis and secretion, by downregulation of, amongst others, SDC4 and RAB27A, two proteins involved in exosome biogenesis. The authors corroborate these results by inducing b-catenin activation using CHIR99021 in a hepatocarcinoma cell line with non-oncogenic bCatenin (Huh7 cells). The authors can further demonstrate convincingly that a reduction in exosome release by hepatocarcinoma spheroids leads to a reduction in immune cell infiltration into the tumor spheroid.

Strengths:

This is a very important and well-conceived study, that appeals to a readership beyond the field of hepatocarcinoma. The authors demonstrate a compelling link between oncogenic bCatenin and exosome biogenesis. Their results are convincing and with well-designed control experiments. The authors included various complementary lines of investigation to verify their findings.

Weaknesses:

One limitation of this study is that the mechanistic relationship of exosome release and how they affect immune cells remains to be elucidated. In this context, the authors conclusions rest on the assumption that hepatocarcinoma immune evasion is based exclusively on the reduced number of exosomes. However, the authors do not analyze exosome composition between exosomes of wild type and oncogenic background, which could be different.

Author response:

The following is the authors’ response to the original reviews.

Reviewer 1:

While the role of Rab27 was strongly examined, the hits of the VAMP proteins were not explored in detail. I was wondering if the decrease in the presence of VAMPS directly suggests the final step of membrane fusion in the exocytosis of EVs is what is being impaired. Or if it is other trafficking steps along the EV secretion pathway.

We appreciate the relevance of this comment and we agree that the decrease of VAMP gene expression in the β-catenin-mutated HepG2 cells could suggest an impairment of the final membrane fusion step in exocytosis of EVs. We have therefore expanded this important point in the discussion (page 10). Indeed, we identified an upregulation of VAMP2, VAMP5 and VAMP8 expressions after mutated β-catenin depletion in the transcriptomic analysis of HepG2 cells. However, these proteins were not detected in the mass spectrometry analysis. Only VAMP3 and VAMP7 proteins were detected in the proteomic analysis without any variation. This is why we didn't focus on this trafficking step, but it could be interesting to explore it further in the future.

Reviewer 2:

(1) In Figure 1F, it is essential to investigate why mass spectrometry analysis indicated no significant changes in SDC4 levels.

We agree with the reviewer that indeed whereas we did observe a significant alteration of syndecan-4 expression at the mRNA level, we did not observe significant changes in syndecan-4 levels by mass spectrometry. One possible explanation is that heparan sulfate proteoglycans like syndecan-4 exhibit a high degree of structural heterogeneity due to the biosynthetic process that produces linear polysaccharides. This characteristic can alter the robustness of mass spectrometry analyses, leading to greater variability.

(2) Figure 2G lacks clarity in explaining how the quantification of MVBs (multivesicular bodies) was conducted.

We apologize for the lack in clarity in explaining how the quantification of MVBs was conducted in figure 2G. The Materials and methods section (part electron microscopy-cells, page 23) has been modified in order to emphasize this point.

(3) In Supplementary Figure 1F, there is a suggestion to highlight exosomes using arrowheads for enhanced clarity.

According to the reviewer’s suggestions, we added arrowheads on supplementary figure 1F in order to highlight the exosomes (page 16). This indeed improves clarity.

(4) Figure 3C prompts a question about the peculiar appearance of Actin staining in KD cells, requiring further investigation.

The peculiar appearance of this intense phalloidin staining between hepatocytes corresponds to bile canaliculi (BC), features of more differentiated HepG2 cells. As phalloidin-stained BC are very bright, this may diminish the visibility of other, thinner actin structures. We decided to change the image of KD cells for a more relevant one (new Figure 3C).

(5) An intriguing avenue for exploration is suggested in testing how the treatment of a GSK inhibitor on HepG2 cells might impact Rab27a and SDC4 expression.

We appreciate the relevance of the suggestion in testing how the treatment of a GSK inhibitor on HepG2 cells might impact Rab27a and SDC4 expression. According to the reviewer’s suggestions, experiments have been carried out and the data are presented in Author response image 1 below. In HepG2 cells, GSK inhibitor stabilized the wild-type β-catenin protein but surprisingly the mutated form of β-catenin is slightly decreased (Author response image 1A). Regarding the expression levels of both Rab27a and SDC4 mRNA, a small increase is observed (Author response image 1B). Rab27a protein is also increased upon the treatment with a GSK inhibitor on HepG2 cells (Author response image 1C). This increased in expression could be due to the decrease of the mutated form of β-catenin in HepG2 cells confirming that Rab27a and SDC4 are repressed by the mutated β-catenin.

Impact of a GSK inhibitor (CHIR99021) on Rab27a and syndecan-4 (SDC4) expressions in HepG2 cells.

HepG2 cells were treated by 3 µM CHIR990221 or DMSO as control for 48h. A) Western-blot (upper panel) and quantification (lower panel) of wild-type (WT) and mutated (MUT) β-catenin proteins in HepG2 cells treated with DMSO (control) or with CHIR990221. B) qRT-PCR analysis of Rab27a and SDC4 expression in HepG2 cells treated with DMSO (control) or with CHIR990221. C) Western-blot (left panel) and quantification (right panel) of Rab27a protein in HepG2 cells treated with DMSO (control) or with CHIR990221. *P<0.05

Reviewer 3:

(1) One limitation of this study is that the mechanistic relationship of exosome release and how they affect immune cells remains to be elucidated. In this context, the authors conclusions rest on the assumption that hepatocarcinoma immune evasion is based exclusively on the reduced number of exosomes. However, the authors do not analyze exosome composition between exosomes of wild type and oncogenic background, which could be different.

We agree that the mechanistic relationship of exosome release and how they affect immune cells remains to be elucidated. In the discussion we mentioned that the content of ß-catenin-regulated EVs remains to be explored to fully understand their function in the immunomodulation of the tumor microenvironment. In this line, we have ongoing experiments in order to analyse the exosomal content in term of proteins and microRNAs. According to our preliminary results, we are able to say that the exosome composition in knock-down mutated ß-catenin HepG2 cells compared to control HepG2 cells seems to be different suggesting not only an involvement of the number of exosomes in the immunomodulation but also of their content.

(2) The manuscript would benefit from minor language editing and the introduction from restructuring to enhance clarity.

The manuscript has now benefited from a language editing thanks to the Professor William A. Thomas (Colby-Sawyer College, New Hampshire). Acknowledgments have been modified (page 12) to thank the Professor William A. Thomas for proof- reading of the manuscript. The introduction has been also restructured and modified according to the reviewer's suggestions to enhance clarity (page 3).

(3) I believe that within the abstract, the authors mean 'defect' not 'default' in the sentence: Then, we demonstrated in 3D spheroid models that activation of β-catenin promotes a decrease of immune cell infiltration through a default in exosome secretion.

We apologize for the mistake between 'default' and 'defect' in the abstract. The abstract has been modified accordingly.

(4) Within the 'Introduction' part of the manuscript, the authors might consider reviewing and reorganizing the first paragraph for more clarity - I suggest leading with the first three sentences of the second paragraph (HCC is the most...) and then introducing b-catenin and the effects and implications of oncogenic ß-catenin in HCC.

If the authors prefer the current structure of the 'Introduction', I would like to propose exchanging some of the wording:

-In line 4: 'despite' instead of 'in front of'? Sentence: Thus, in front of the therapeutic revolution for cancers, with the emergence of immunotherapy and more particularly immune checkpoint inhibitors (anti-PD1, anti-PD-L1)

-Additionally in line 7: In these tumors, the oncogenic β-catenin is able to set up a microenvironment that favors tumor progression notably by promoting immune escape. Here, 'establish' might be a better choice instead of 'set up' - In line 9 I suggest rephrasing the sentence: Few studies have reported that the defect of intercellular communication between cancer cells and immune cells is partly mediated by a decrease of chemokines production leading to a reduction of immune infiltrates.... and maybe adding a reference here.

The introduction has been altered accordingly. Thanks for these suggestions that helped us to improve our manuscript.

Version 2

Abstract

Abstract

Immune checkpoint inhibitors have produced encouraging results in cancer patients. However, the majority of β-catenin-mutated tumors have been described as lacking immune infiltrates and resistant to immunotherapy. The mechanisms by which oncogenic β-catenin affects immune surveillance remain unclear. Herein, we highlighted the involvement of β-catenin in the regulation of the exosomal pathway and, by extension, in immune/cancer cell communication in hepatocellular carcinoma (HCC). We showed that mutated β-catenin represses expression of SDC4 and RAB27A, two main actors in exosome biogenesis, in both liver cancer cell lines and HCC patient samples. Using nanoparticle tracking analysis and live-cell imaging, we further demonstrated that activated β-catenin represses exosome release. Then, we demonstrated in 3D spheroid models that activation of β-catenin promotes a decrease in immune cell infiltration through a defect in exosome secretion. Taken together, our results provide the first evidence that oncogenic β-catenin plays a key role in exosome biogenesis. Our study gives new insight into the impact of β-catenin mutations on tumor microenvironment remodeling, which could lead to the development of new strategies to enhance immunotherapeutic response.

Introduction

Hepatocellular carcinoma (HCC) is the most frequent form of primary liver adult cancer (80% of cases). HCC represents the 6th most common diagnosed cancer worldwide and the 3rd leading cause of cancer-related death (1). This pathology has a poor prognosis and the diagnosis is often made at an advanced stage for which treatment options are limited, especially if surgery is no longer possible. In 2020, the combination of Atezolizumab (anti-PD-L1) and Bevacizumab (anti-VEGF) became the first-line FDA-approved therapy for advanced HCC (2). Despite this therapeutic advance, several studies emphasized that β-catenin-mutated HCC are devoid of immune infiltrates and are resistant to immunotherapy (3–6). CTNNB1 gene mutations are found in 30-40% of HCC and trigger uncontrolled transcriptional activity (7,8). These mutations prevent β-catenin degradation, fostering its role as a transcriptional co-factor and thus the expression of genes involved in cell proliferation and invasion. This altered pattern of gene expression could provide a key to understanding to the observed immune evasion. β-catenin, encoded by the CTNNB1 gene, is a main oncogene, mutated in various cancers, such as melanoma and liver, endometrial, and colorectal cancers (9). These β-catenin-mutated tumors share specific features, prominent among them a microenvironment devoid of immune infiltrates. Thus, despite of a therapeutic revolution for cancer treatment, with the emergence of immunotherapy and more particularly immune checkpoint inhibitors (anti-PD1, anti-PD-L1), most of the β-catenin-mutated tumors remain resistant to immunotherapy (10–15). In these tumors, the oncogenic β-catenin is able to establish a microenvironment that favors tumor progression notably by promoting immune escape. Few studies have reported that the oncogenic β-catenin is implicated in the impairment of intercellular communication between cancer cells and immune cells, partly through soluble molecules such as cytokines.. In melanoma and HCC, a decrease in CCL4 and CCL5, leads to a defective recruitment of dendritic cells and consequently impaired T-cell activity and immune escape (5,11). As immunotherapies are now part of the growing arsenal for treating cancers, it is important to better understand the mechanisms underlying this immune escape phenotype. In this study we focused on β-catenin-mutated HCC and intercellular communication through extracellular vesicles.

Tumor-derived extracellular vesicles (EVs) are increasingly described as important actors in tumor microenvironment communication (16); but no studies explored their role in the immune escape and immunotherapy resistance of β-catenin-mutated tumors. EVs are nanometric phospholipid bilayer transport vesicles, which contain various cargoes such as RNAs or proteins and that impact the properties and functions of recipient cells. EVs can be classified into two main categories: microvesicles and exosomes. Microvesicles (50 to 1000 nm) correspond to EVs directly secreted from the plasma membrane (PM). By contrats, exosomes (50 to 150 nm) are released from an endosomal compartment, the multivesicular body (MVB), by its fusion with the cell surface (16). Among the actors implicated in exosome biogenesis and secretion, syndecan-4, involved in the endosomal membrane budding (17) and Rab27a, which promotes the docking of MVB to the PM, play key roles in exosome release (18). As important mediators of cell-to-cell communication, EVs can regulate tumor growth, angiogenesis, invasion and infiltration of immune cells into the tumor (19). In the context of HCC, it was reported that EVs derived from tumor cells can modulate epithelial-to-mesenchymal transition, intrahepatic metastasis and tumor immunity (20).

Based on these considerations, the current study investigated the role of exosomes in mutated β-catenin-mediated immune evasion in HCC. We demonstrated that mutated β-catenin represses syndecan-4 and Rab27a expression, thus altering exosomal secretion from HCC cells, in turn leading to a defective recruitment of immune cells in tumors.

Results

Silencing of mutated β-catenin increases exosome biogenesis-associated gene expression in HepG2 cells

CTNNB1 mutations in HCC are often monoallelic, leaving a wild-type allele in tumor cells. To study the oncogenic β-catenin specifically, we made use of the dual β- catenin knockdown (KD) HepG2 model we published recently (21). The transcriptional analysis of HepG2 KD cells revealed that the expression of 1973 genes (log2(Fold-change) >0.5 and <-0.5) is modulated upon mutated β-catenin silencing (21) (Figure 1a). Gene ontology analysis (FunRich) performed on upregulated genes showed that mutated β-catenin silencing enhanced significantly the expression of genes linked to specific cellular components, including exosomes (Figure 1b). Moreover, silencing of the mutated β-catenin led to the overexpression of 17 genes associated with exosome biogenesis, such as RAB27A and SDC4 (Figure 1c). Similarly, proteomic analysis of HepG2 KD cells revealed exosomes as the main cellular component displaying up-regulated proteins upon mutated β- catenin silencing (Figures 1d-1e). Consistent with the transcriptomic analysis, we found Rab27a and Syndecan-4 among up-regulated proteins, results (Figure 1f), even if the variability was found for Syndecan-4 protein levels. These results suggest the involvement of mutated β-catenin in the regulation of the exosomal pathway.

Mutated β-catenin regulates exosome biogenesis gene expression in HepG2 cells.

(a-f) HepG2 cells were transfected with siRNAs targeting mutated β- catenin or control siRNAs. (a) Volcano plot of deregulated genes identified by microarray based-transcriptomic analysis. Red and blue dots indicated respectively significantly up- and down-regulated genes. (b) Upregulation of cellular component genes using FunRich software. (c) Upregulated genes associated with exosome biogenesis pathway. The graph indicates the fold change (FC) when comparing the mutated β-catenin silencing condition with the control condition. (d) Volcano plot of deregulated proteins identified by mass spectrometry. Red and blue dots indicated respectively significantly up- and down-regulated proteins. (e) Upregulation of cellular component proteins identified using FunRich software. (f) Upregulated proteins associated with exosome biogenesis pathway. The graph indicates the fold change when comparing the mutated β-catenin silencing condition with the control condition. Results are expressed as Mean ± SEM, two-tailed Student’s t-test analysis. *P<0.05;**P<0.01; ***P<0.001; ****P<0.0001.

Stable depletion of mutated β-catenin increases exosome secretion in the HepG2 cell model

In the model presented above, siRNAs were used to transiently silence mutated β- catenin expression in HepG2 cells (21). To improve this method, the same sequences were used to develop an inducible shRNA strategy to stably inhibit the expression of the mutated β-catenin. Through the use of doxycycline-inducible promoters we were able to regulate the expression of shRNAs over time for more in- depth and specific investigations of β-catenin function in various cellular processes. In HepG2 cells, this strategy reduced protein expression by 62% for mutated β- catenin and by 42% for CyclinD1, a β-catenin-positive transcriptional target (Figure 2a). The model was also validated regarding mRNA expression of several β-catenin transcriptional targets. We noted a decrease in the expression of CCND1 and AXIN2 (positive targets) and an increase in the expression of ARG1 expression (negative target) (Figure sup 1a). Also, as previously described with the siRNA strategy (21), after mutated β-catenin depletion in HepG2 cells, we observed an increase in the number and size of bile canaliculi (BC), feature of more differentiated cells (Figures sup 1b-1c). Nanoparticle-tracking analysis (NTA) revealed that depleting mutated β- catenin using either siRNA or shRNA increased the number of secreted particles (Figure 2b, Figure sup 1d). The particle mean size (100-150 nm), unaltered by either approach, was consistent with the size of typical small EVs, such as exosomes (Figure 2b, Figure sup 1d). To analyze the exosomal release in living HepG2 cells, we used a pH-sensitive reporter (CD63-pHluorin) that allowed the visualization of MVB–PM fusion by total internal reflection fluorescence (TIRF) microscopy (22). The tetraspanin CD63 is a well-known exosome surface marker (23,24). In cells expressing CD63-pHluorin, we could detect multiple discrete increases in the fluorescence signal at the PM over the time suggesting ongoing fusion of CD63-pHluorin–positive acidic vesicles with the PM (Figure sup 1e). We found that mutated β-catenin depletion in HepG2 cells significantly increased the number of fusion events associated with secreted exosomes (Figure 2c). EVs released by HepG2 cells were next isolated by differential ultracentrifugation and characterized (Figure 2d). First, western-blot analysis of isolated EVs confirmed the expression of the CD63 exosomal marker. It is noteworthy that we observed an increase in CD63 expression after mutated β-catenin depletion, suggesting an overall increase in the exosomal fraction (Figure 2e). This increased expression of CD63 was also confirmed with the proteomic analysis of HepG2 KD cells (Figure 1f). Moreover, the WT and the mutated forms of the β-catenin protein were also found in HepG2-derived EVs, and the expression of mutated β-catenin was decreased relative to the cells of origin (Figure 2e). Using transmission electron microscopy (TEM), we demonstrated that HepG2- derived EVs have the morphology (flat cup-shaped) and size (50-150nm) of typical (25) (Figures 2f and sup 1f). To better understand the regulation of exosome secretion by mutated β-catenin, we next used TEM to visualize MVBs, the origin of exosomes. We observed a higher number of MVBs per cell and an increase in their diameter upon removal of mutated β-catenin (Figure 2g). Taken together these results highlight that mutated β-catenin represses exosome secretion.

Mutated β-catenin controls exosome secretion in HepG2 cells.

(a-g) HepG2-shβCat MUT and HepG2-shCtrl cells were treated with doxycycline (DOX) to silence or not mutated β-catenin. (a) Analysis of β-catenin and CyclinD1 expression by western-blot. Stain free was used as loading control. Graphs show the quantification of seven independent experiments. (b) Nanoparticle tracking analysis of supernatant. Graphs show the quantification of seven independent experiments. (c) Quantification of CD63-pHluorin MVB–PM fusion events visualized by live TIRF microscopy. Depicted data are representative of three independent experiments, each dot represents one cell. Images represent the cell mask (white) and red dots corresponding to fusion events. Scale bar: 10µm. (d) Extracellular vesicles (EVs) isolation protocol. Created with BioRender.com. (e) Analysis of β-catenin and CD63 expression in HepG2-derived EVs. Stain free was used as loading control. Graphs show the quantification of four independent experiments. (f) Transmission electron microscopy images of HepG2-derived EVs by close-up. Scale bar: 100nm. The graph shows the diameter quantification of EVs (n=93). (g) Electron microscopy images of HepG2 shCtrl and shβcat MUT cells showing multivesicular bodies (MVBs) (yellow arrowheads). Scale bar: 2 µm (zoom: 500 nm). The graphs show the quantification of the number of MVBs per cell and the MVB diameter. (a-f) Results are expressed as Mean ± SEM, two-tailed Student’s t-test analysis. *P<0.05; **P<0.01; ***P<0.001; ns, non-significant.

Mutated β-catenin decreases Syndecan-4 and Rab27a expression in liver cancer cell lines

To explore the molecular aspect of the alteration of exosome secretion, we first attempted to confirm the data obtained in our transcriptomic and proteomic approaches. In HepG2 cells, we confirmed that transient (Figures sup 2a-b) and stable (Figures 3a-3b) mutated β-catenin depletion increased gene and protein expression of both SDC4 and RAB27A. Due to the lack of efficient antibodies against SDC4, we were unable to analyze SDC4 protein expression by Western-Blot. However, we were able to confirm the increased expression of Syndecan-4 and Rab27a by immunofluorescence (Figures 3c-3d). We then extended these results to other liver cancer cell lines. We analyzed SDC4, RAB27A, ARG1 and AXIN2 basal mRNA expression in five different liver cancer cell lines, each with a different CTNNB1 mutational status, and identified positive correlations between SDC4 and RAB27A with ARG1 (β-catenin-negative target) and negative correlations between SDC4 and RAB27A with AXIN2 (β-catenin-positive target) (Figure sup 2c-d). Upon analyzing the basal expression of Rab27a protein in these five cell lines, we identified a negative correlation between β-catenin and Rab27a expression (Figures sup 2f- 2g). In the two liver cancer cell lines bearing a β-catenin point mutation (Huh6 and SNU398 cells), inducible shRNA strategy with doxycycline reduced β-catenin protein expression by 63% and 49%, respectively (Figures sup 2h-2i). Associated with that reduction in β-catenin protein expression we observed a decrease in CCND1 and AXIN2 mRNA expression and an increase in ARG1 mRNA expression in both cell lines (Figures sup 2j-2k). As in HepG2 cells, β-catenin depletion also increased SDC4 and RAB27A gene expression and Rab27a protein expression in Huh6 and SNU398 cells treated with doxycycline (Figures 3e-h). In order to mimic the β-catenin activation in a non-mutated HCC cell line, we treated Huh7 cells with a GSK3 inhibitor (CHIR99021) limiting β-catenin phosphorylation and degradation. This CHIR99021 treatment increased β-catenin protein expression (2.37-fold) (Figure 3i) and induced its translocation into the nucleus in Huh7 cells (Figure 3j) where concomitant deregulation of β-catenin transcriptional targets was observed (Figure sup 2l). Consistent with our earlier data using HepG2 cells, CHIR99021 treatment decreased SDC4 and RAB27A gene expression, and Rab27a protein level in Huh7 cells (Figures 3k-l). Then, we used TIRF microscopy to analyze exosomal release in living Huh7 cells expressing CD63-pHluorin. We found that CHIR99021 treatment significantly decreased the number of fusion events in Huh7 cells (Figure 3m). All together, these results revealed that β-catenin mutation/activation represses the expression of syndecan-4 and Rab27A, two proteins involved in exosome biogenesis, in association with repression of exosomal secretion.

Activated β-catenin represses SDC4 and RAB27A expression in liver cancer cells.

(a-d) HepG2 cells were treated with doxycycline (DOX) to express either a control shRNA (shCtrl) or a shRNA targeting mutated β-catenin (shβcat MUT). (a) Analysis of SDC4 and RAB27A mRNA expression by qRT-PCR. Graphs show the quantification. (b) Analysis of Rab27a protein expression by western-blot. Stain free was used as loading control. The graph shows the quantification. (c-d) Epifluorescence images of HepG2 shCtrl and shβcat MUT cells stained with Syndecan-4 and Rab27a antibodies (red), Phalloidin (green), Hoechst (blue). Scale bar: 20 µm. Graphs show the quantification of the fluorescence intensity per image and divided per nuclei. Depicted data are representative of three independent experiments, each dot represents one image. (e) Analysis of SDC4 and RAB27A mRNA expression in Huh6 cells expressing either a control shRNA (shCtrl) or a shRNA targeting β-catenin (shβcat) treated with DOX. The graph shows the quantification. (f) Analysis of SDC4 and RAB27A mRNA expression in SNU398 cells expressing either a control shRNA (shCtrl) or a shRNA targeting β-catenin (shβcat) treated with DOX. The graph shows the quantification. (g-h) Analysis of Rab27a protein expression in Huh6 (g) or SNU398 (h) cells expressing either a control shRNA (shCtrl) or a shRNA targeting β-catenin (shβcat). Stain free was used as loading control. The graphs show the quantification. (i-m) Huh7 cells treated with DMSO or CHIR99021 (3µM) for 48 hours. (i) Analysis of β-catenin expression by western-blot. The graph shows the quantification. (j) Epifluorescence images of cells stained with β-catenin antibody (red), Phalloidin (green), Hoechst (blue). Scale bar: 20 µm. (k) Analysis of SDC4 and RAB27A mRNA expression by qRT-PCR. The graphs show the quantification. (l) Analysis of Rab27a protein expression by western-blot. The graph shows the quantification. (m) Quantification of CD63- pHluorin MVB–PM fusion events visualized by live TIRF microscopy. Depicted data are representative of three independent experiments; each dot represents one cell. Images represent the cell mask (white) and red dots corresponding to fusion events. (a-m) All graphs show the quantification of at least three independent experiments. Results are expressed as Mean ± SEM, two-tailed Student’s t-test analysis. *P<0.05; **P<0.01; ***P<0.001.

Mutated β-catenin promotes immune evasion in HCC cells through exosomes

Our next goal was to investigate the possible role of mutated β-catenin in HCC immune escape. We used 3D spheroid models from two liver cancer cell lines (HepG2 and Huh7) and after 24h co-culture of spheroids with Peripheral Blood Mononuclear Cells (PBMC), immune cell infiltration was evaluated by flow cytometry (Figures 4a-4b). First of all, we assessed the capability of PBMC to invade spheroids formed from Huh7 cells treated with CHIR9901. We observed a decrease in PBMC infiltration and an increase in tumor cell survival in Huh7 cells were treated with CHIR9901 compared to the results seen in the control condition (Figure 4c). Of note, we confirmed that CHIR99021 increased β-catenin and decreased Rab27a protein expression in Huh7 spheroids with no impact on the cell viability (Figures sup 3a-3b). Next, we showed that the depletion of mutated β-catenin protein enhanced PBMC infiltration in HepG2 cells associated with a decrease of tumor cell survival (Figure 4d). We verified that mutated β-catenin protein was correctly depleted in HepG2 spheroids, with an increase in Rab27a protein expression and no impact on cell viability (Figures sup 3c-d). Altogether, these results confirm the involvement of mutated β-catenin in immune evasion. We next sought to investigate the role of exosomes in this process. In order to blunt exosome secretion, we used siRNA directed against Rab27a in spheroids formed with HepG2 cells depleted of mutated β-catenin. Rab27a depletion decreased the number of secreted particles without any effect on cell viability in HepG2 shβcat MUT spheroids (Figures sup 3e-g). Furthermore, depletion of Rab27a decreased PBMC infiltration associated with an increase of tumor cell survival (Figure 4e). Taken together, these results strongly suggest a determinant role for exosomes in immune cell infiltration.

Activated β-catenin represses immune infiltration in liver cancer spheroids through exosomes. (a)

Peripheral Blood Mononuclear Cells (PBMC) infiltration analysis protocol. Created with BioRender.com. (b) Images of HepG2 spheroids expressing a control shRNA incubated or not with PBMC for 24 hours. (c) Analysis of PBMC infiltration and tumor cell survival in Huh7 spheroids treated with DMSO or CHIR99021. (d) Analysis of PBMC infiltration and tumor cell survival in HepG2 spheroids expressing control shRNA (shCtrl) or shRNA targeting mutated β- catenin (shβcat MUT). (e) Analysis of PBMC infiltration and tumor cell survival in HepG2 spheroids co-expressing shRNA targeting mutated β-catenin (shβcat MUT) and control siRNA (siCtrl) or siRNA targeting Rab27A (siRab27A). (c-e) Graphs show the quantification of six independent experiments. Results are expressed as Mean ± SEM, one or two-tailed Student’s t-test analysis. *P<0.05; **P<0.01; ***P<0.001.

CTNNB1 mutations are associated with low expression of exosomal biogenesis-associated genes in HCC patient samples

To determine whether markers of exosomal biogenesis are deregulated in HCC human samples, we assessed the expression of SDC4 and RAB27A using public transcriptomic data from two different cohorts (Figure 5a: TGCA data sets from Cbioportal, Figure 5b: Boyault et al. article (26)). In both set of data, we showed that SDC4 and RAB27A gene expressions were significantly reduced in CTNNB1- mutated HCCs compare to non-mutated tumors (Figures 5a-5b). Furthermore, Pearson’s correlation analysis of these two cohorts revealed that RAB27A and SDC4 gene expression were positively correlated with the expression of ARG1 and PCK1 (β-catenin-negative targets) (27) and negatively correlated with the expression of GLUL, CCND1, AXIN2, LGR5, FAT1 and BMP4 (β-catenin-positive targets) (Figures sup 4a-b). We then performed immunohistochemical analysis on 56 human HCC samples with or without CTNNB1 mutations (Figure sup 4c). We found that tumors strongly positive for glutamine synthetase (a well-known β-catenin-positive target in the liver (28)), were correlated with low levels of Rab27a protein expression (Figure 5c), supporting the link between β-catenin activation and regulation of exosomal secretion through a decrease in Rab27a expression.

Downregulation of SDC4 and RAB27A in human HCC samples with CTNNB1 mutations. (a)

Analysis of SDC4 and RAB27A mRNA expression in β- catenin mutated (red) and non-mutated (blue) HCC (Cbioportal cohort, n=366). (b) Analysis of SDC4 and RAB27A mRNA expression in β-catenin mutated (red) and non-mutated (blue) HCC (Boyault et al. cohort, n=56). (c) Immunohistochemistry (IHC) analysis of glutamine synthetase (GS) and Rab27a in HCC samples presenting or not β-catenin mutations. Scale bar: 1mm. (T: tumoral, NT: non tumoral). The upper graphs show the quantification of the percentage of cells positive for GS or Rab27A. The analysis was split into two cohorts (circle and triangle). The lower graph shows the Pearson correlation between Rab27a and GS protein expression (n=56). Results are expressed as Mean ± SEM, two-tailed Student’s t-test analysis or Pearson correlation test. *P<0.05; **P<0.01; ***P<0.001; ****P<0.0001.

Discussion

β-catenin is a main oncogene in tumors (9) and several studies have reported a key role for mutated β-catenin in immune escape in several types of cancer such as HCC, colorectal cancer, melanoma and glioblastoma (6,11,12,15), but its mechanisms of action in this process remain to be elucidated. During the past decades, β-catenin mutated HCC have been classified at the molecular, histo- pathological and clinical levels as liver cancers with specific features (29–31). More recently, these tumors have been described as cold with low immune infiltrates and resistant to immunotherapy (3–6).

In this study we deciphered the role of mutated β-catenin in cell communication between cancer cells and immune cells. Based on omic results, we hypothesized a role for EV in the control of immune infiltrates in β-catenin-mutated HCC. Following ISEV guidelines (32), we applied various techniques to address the impact of β- catenin activation on EV secretion. Using NTA, we demonstrated that HepG2 cells silenced for mutated β-catenin secrete more particles than do control cells. Our different analyses and results (TIRF with the CD63 exosomal marker, TEM for the EVs shape and size, regulation of Rab27a expression) led us to take a strong interest in small EVs (i.e. exosomes), a subpopulation of EVs. Using real-time visualization of the release of CD63-enriched EVs we demonstrated that mutated β- catenin silencing in HepG2 cells favors small EV secretion. The link between activation of β-catenin and secretion of small EVs was further confirmed in β-catenin non-mutated cells, Huh7 cells, treated with GSK3 inhibitor, suggesting that activation of the Wnt/β-catenin pathway may result in a repression of the EV machinery in tumor cells. This link between β-catenin and EVs was poorly explored in the literature and essentially restricted to the identification of β-catenin as a common EV cargo (33). For example, it was shown that tetraspanins expression reduces the cellular pool of β-catenin by enhancing the exosome-associated export of β-catenin from the cell and that this exosomal discharge of β-catenin downregulates the Wnt signaling pathway (34). More recently, using a genome-wide CRISPR/Cas9-mediated screening to identify genes involved in EV release, Lu et al. identified the Wnt signaling pathway in the top hits (35). Indeed, by this approach, a number of Wnt- signaling modulators appeared among the candidates as EV biogenesis regulators, which were further validated in chronic myelogenous leukemia K562 cells. Lu et al. also demonstrated that Wnt-mediated GSK3 inactivation, in K562 cells, regulated EV release in a Rab27a dependent manner (35). Herein, we added new findings reporting that this regulation also occurs in solid tumors. Moreover, we demonstrated for the first time the specific involvement of the oncogenic β-catenin in the regulation of exosome biogenesis.

At the molecular level, thanks to global approaches we found that oncogenic β- catenin represses the expression of genes involved in the EV machinery. We indeed found that genes from the Rab (RAB8B, RAB11A, RAB11B, RAB17, RAB27A), Vamp (VAMP2, VAMP5, VAMP8) and Vps (VPS4A, VPS4B, VPS16, VPS28) families were upregulated upon removal of the oncogenic β-catenin in HepG2 cells. Therefore, several trafficking steps could be impaired along the EV secretion pathway, the MVB biogenesis with the Vps, the transport and docking with the Rab and the final step of membrane fusion with the Vamp (16). Herein, we focused on RAB27A and SDC4 genes, which were the most dysregulated in our transcriptomic analysis. We showed that mutated β-catenin represses their gene expression in liver cancer cell lines. In non-mutated β-catenin HCC cells, activation of β-catenin showed that both gene and protein expression of Rab27a were decreased. We further found a correlation between the presence of CTNNB1 mutations and the low expression of RAB27A and SDC4 in human HCC patient samples. The way RAB27A and SDC4 gene expression is repressed by β-catenin remains to be explored. We described positive correlation between expression of SDC4 and RAB27A and expression of negative transcriptional targets of β-catenin, as well as a negative correlation with expression of positive targets in tumors. Even if the mechanism by which β-catenin and LEF/TCF transcription factors led to target gene expression activation is well known (21,36,37), it is not the case for the repression process. In melanoma the transcription repressor ATF3 induced by the Wnt/β-catenin pathway has been associated with CCL4 repression (11). Analysis of SDC4 and RAB27A promoter sequences revealed the presence of consensus sequences of Lef1, Tcf4 and Atf3 binding (data not shown), suggesting that these newly identified target genes could be regulated at the transcriptional level by transcription factors known to associate with β-catenin. In this line, SDC4 was found to be a negative Tcf4 target gene in HCC mouse models (38).

Following the EV machinery alteration, we further addressed the impact of the β- catenin-mediated repression of EV secretion on immune cell infiltration in tumors. Herein, the functional role of these tumor-derived exosomes in tumor immunity was studied using liver cancer 3D models and PBMC. Using siRNA depletion of Rab27a, we showed that mutated β-catenin promotes a decrease of immune cell infiltration in micro-tumors through a decrease of exosome secretion by . The small GTPase Rab27a has been implicated in several cancer progression mechanisms such as metastasis, cell invasion or migration (39). The oncogenic function of Rab27a is mainly due to its function as regulator of exosome secretion, which modulates cancer cell function and tumor microenvironment. In mammary carcinoma cells, the decrease of exosome secretion by Rab27a inhibition resulted in a diminution of primary tumor growth and lung metastatic dissemination (40). However, shRNA knockdown of Rab27a in a prostate cancer model revealed a dominant immunosuppressive role for EVs during antigen cross presentation (41). These data are inconsistent with our own, meaning so it seems that Rab27a-induced exosomal secretion could be a “double-edged sword” in tumor progression. This could be due to the content of EVs. In fact, EVs are known to contain heterogeneous cargos that can play a critical role in immunomodulation (42). For example, the main immunosuppressive cytokine TGF-β expressed by tumoral-derived EVs has been identified in Natural Killer cell function suppression (43,44). EVs released by tumoral cells can also carry several miRNAs mainly known to be involved in immune suppression (42). MiR-103a and Let-7a in tumor EVs promote M2-like polarization in lung and melanoma cancers (45,46). Interestingly, immunomodulatory proteins such as CSF-1, CCL2, FTH, FTL, and TGF-β chemokines have been identified in tumor exosomes (46). The content of β-catenin-regulated EVs remains to be explored to fully understand their function in the immunomodulation of the tumor microenvironment.

In conclusion, this study is the first description of the oncogenic β-catenin as a key regulator of the exosome machinery. Moreover, it gave new insights regarding the regulation of tumor-derived exosome production by the β-catenin pathway and their involvement in the immune escape observed in β-catenin mutated HCC. Given that β-catenin is also mutated in melanoma, endometrial and colorectal cancers, these new findings may have a broader significance. Our data indicate that exosomes could be a promising biomarker and that restoring the tumor exosomal secretion of β- catenin-mutated tumors may increase the communication between tumor and immune cells, and as a result, turn cold tumors hot.

Supporting information